Interpretable Machine Learning

### 2019-01-10 chunks/story-wines.Rmd
continues talk

TODO: Add wine csv and R script

@@

# What is Machine Learning?

<div class="notes">
Machine learning is a method for teaching computers to make and improve predictions or behaviours based on data.
</div>

## {data-background=../images/magic.jpg data-background-size=contain}

@@

Red and white variants of the Portuguese "Vinho Verde" wine.

@@

TODO: Draw image with features.

1 - fixed acidity
2 - volatile acidity
3 - citric acid
4 - residual sugar
5 - chlorides
6 - free sulfur dioxide
7 - total sulfur dioxide
8 - density
9 - pH
10 - sulphates
11 - alcohol
Output variable (based on sensory data):
12 - quality (score between 0 and 10)

Draw:
Volatile acidity as polish remover, pH as pH paper, alchohol as liqour bottle, chlorides as salt prinkler, citric acid as lemon, 
sulfur as mushroom?,

@@

```{r feature-effects}
effs = FeatureEffects$new(pred)
plot(effs, ncols = 4)
```

```{r feature-effect-alcohol}
eff = FeatureEffect$new(pred, "alcohol")
plot(eff)
```

## Rule of thumb what makes wine good?

```{r surrogate}
sbrl()
tree = TreeSurrogate$new(pred, maxdepth  = 2)
plot(tree)
```

## Customers want to have individual explanations

TODO: Image where customers checks wine?

@@

@@

TODO: Draw features as players and within coalitions

```{r shapley}
shap = Shapley$new(pred, x.interest = worst)
plot(shap)
```

## Volatile Acidity

```{r volatile-acidity}
eff = FeatureEffect$new(pred, "volatile.acidity", method = "pdp+ice")
plot(eff)
```
## Volatile Acidity

‘Although the presence of high amounts of VA is considered undesirable, in some cases a touch of volatility is no bad thing,’ said Natasha Hughes MW in her guide to wine ‘flaws’.

```{r volatile-acidity}

# Get wines with low volatile acidity
predictions = pred$predict(wine_subset)

wine2 = wine_subset
wine2$volatile.acidity = wine2$volatile.acidity + 0.2
predictions2 = pred$predict(wine2)

plot(hist(predictions2$.prediction - predictions$.prediction))

wine_improve = which(predictions2$.prediction - predictions$.prediction > 0.3)

eff = FeatureEffect$new(pred, "volatile.acidity", method = "ice")

res = eff$results
res$improves = res$.id %in% wine_improve

ggplot(res) + 
  geom_line(aes(x = volatile.acidity, y = .y.hat, group = .id), alpha = 0.1) + 
  facet_wrap("improves")
```

## LIME

TODO: Explain with LIME

## Sulfur

At low levels reduction causes truffle, mushroom, radish, and green olive-like smells to wine. These have been accepted as positive traits and increase complexity in the wine. At higher levels, reduction is turns into a fault, giving wines a pungent smell of rotten egg, onion, garlic, cooked cabbage, canned corn, or even burnt rubber.

```{r sulfur}
```

>>>>

@@

@@

Free online dataset about red and white variants of the Portuguese "Vinho Verde" wine.

@@

- fixed acidity
- volatile acidity
- citric acid
- residual sugar
- chlorides
- free sulfur dioxide
- total sulfur dioxide
- density
- pH
- sulphates
- alcohol
- quality based on sensory data (0-10)

<!-- Draw: -->
<!-- Volatile acidity as polish remover, pH as pH paper, alchohol as liqour bottle, chlorides as salt prinkler, citric acid as lemon,  -->
<!-- sulfur as mushroom?,  -->

@@

```{r}
effs = FeatureEffects$new(pred)
plot(effs, ncols = 4)
```

## How do features affect predictions?
Method: Partial dependence plot, individual conditional expectation curves
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'pdp+ice')
plot(eff)
```

## Interactions with type of wine?
Method: H-statistic for interactions

```{r}
inter = Interaction$new(pred)
plot(inter)
```

## Interactions of alcohol?
Method: H-statistic for interactions

```{r}
inter.a = Interaction$new(pred, "alcohol")
plot(inter.a)
```

## Rule of thumb for wine quality?
Method: Surrogate Model

```{r surrogate}
tree = TreeSurrogate$new(pred, maxdepth  = 2)
plot(tree$tree)
```

@@

A customer gets a really bad prediction. What was the reason?

@@

```{r}
shap = Shapley$new(pred, x.interest = worst)
plot(shap)
```

## The best wine

```{r best-wine}
# find bad wine in data
max_pred = max(predictions)
best = wine[which(predictions == max_pred),]

# explain with shapley
knitr::kable(t(best))
# explain with cf?

max_pred

## The best wine
Technique: Shapley Values

```{r}
shap = Shapley$new(pred, x.interest = best)
plot(shap)
```
